---
title: "sarima"
author: "Yutao Chen"
date: "2023-04-13"
output: html_document
---

```{r, echo=FALSE,message=FALSE,warning=FALSE}
library(tidyverse)
library(ggplot2)
library(forecast)
library(astsa) 
library(xts)
library(tseries)
library(fpp2)
library(fma)
library(lubridate)
library(tidyverse)
library(TSstudio)
library(quantmod)
library(tidyquant)
library(plotly)
library(ggplot2)
```

#  Introduction

Load the Climate data `climateDC.csv` that you used for the Quiz.

Downloaded the dataset using NOAA.

<https://www.ncdc.noaa.gov/cdo-web/search>

Here, I'm pre-processing the data.

```{r,message=FALSE,warning=FALSE}
data<-read.csv("climateDC.csv",header =TRUE)
head(data)

#########################
### Check for missing values ########

# is.na(data)

#Extract rows with missing values
df_na_rows <- data[which(rowSums(is.na(data)) > 0),]

#Extract columns with missing values
df_na_cols <- data[, which(colSums(is.na(data)) > 0)]

library(imputeTS)
imputed_time_series <- na_ma(data, k = 4, weighting = "exponential")

```

```{r,message=FALSE,warning=FALSE}
df<-data.frame(imputed_time_series)
str(df)
df$DATE<-as.Date(df$DATE,format = "%m/%d/%y")

library(lubridate)
df$year <- year(df$DATE)
head(df)


######## getting data for only one station
table(df$NAME)
newdf<-df[df$NAME=="NATIONAL ARBORETUM DC, MD US",]
head(newdf)

############ for a larger window
#library(dplyr)
start_date <- as.Date("2012-01-01")
data_subset <- filter(newdf, DATE >= start_date)

##### Just looking at some Variables ###########
#fig <- plot_ly(data_subset, x = ~DATE, y = ~TOBS,name = 'Temperature', type = 'scatter', mode = 'lines',size = ~TOBS)
#fig <- fig %>% add_trace(y = ~SNOW, size = ~SNOW,name = 'Snow Fall', mode = 'markers')
#fig <- fig %>% add_trace(y = ~SNWD, size = ~PRCP,name = 'Percipitation', mode = 'markers')

#fig <- fig %>% layout(title = 'Weather Data in Washington DC',
#                      xaxis = list(showgrid = FALSE),
#                      yaxis = list(showgrid = FALSE))

#fig
write.csv(newdf, "DC.csv")
```


**Here I'm creating time series object for the original data for temperature**

`temp` is the original data for temperature.Whereas `temp_month` is data that I aggregated over month so I get monthly temperature.

```{r,message=FALSE,warning=FALSE}
temp<-ts(newdf$TOBS,star=decimal_date(as.Date("1990-01-01",format = "%Y-%m-%d")),frequency = 12)
 

### can get monthly data
# Get mean value for each month
mean_data <- newdf %>% 
  mutate(month = month(DATE), year = year(DATE)) %>% 
  group_by(year, month) %>% 
  summarize(mean_value = mean(TOBS))

temp_month<-ts(mean_data$mean_value,star=decimal_date(as.Date("1973-01-01",format = "%Y-%m-%d")),frequency = 12)
  


```


**Here I'm creating time series object for a smaller window of the data for temperature**

`tempL` is the for temperature for this data in a smaller window. Similarly, `temp_monthL` is data that I aggregated over month so I get monthly temperature.


```{r,message=FALSE,warning=FALSE}
########### for a smaller window ###################

tempL<-ts(data_subset$TOBS,star=decimal_date(as.Date("2012-01-01",format = "%Y-%m-%d")),frequency = 365.25)

### can get monthly data
# Get mean value for each month
mean_data <- data_subset  %>% 
  mutate(month = month(DATE), year = year(DATE)) %>% 
  group_by(year, month) %>% 
  summarize(mean_value = mean(TOBS))


temp_monthL<<-ts(mean_data$mean_value,star=decimal_date(as.Date("2012-01-01",format = "%Y-%m-%d")),frequency = 12)


```

```{r}
plot(temp_month, xlab="DATE", ylab="Temperature", main="Plot for Monthly Temperature in DC, From 1973")
```

```{r}
plot(decompose(temp_month, type = "additive"))
Acf(temp_month,50, main="ACF Plot of Temperature")
Pacf(temp_month,50, main="PACF Plot of Temperature")
```

```{r}
# Differencing plots
temp_month %>% diff() %>% ggtsdisplay(main = "First Order Differencing")
temp_month %>% diff(lag=12) %>% ggtsdisplay(main = "First Seasonal Differencing")
temp_month %>% diff(lag=12) %>% diff() %>% ggtsdisplay(main = "Do Both")
d1 <- temp_month %>% diff() %>% diff(lag=12)
```

```{r, warning=FALSE}
tseries::adf.test(d1)
```

```{r}

#write a funtion
SARIMA.c=function(p1,p2,q1,q2,P1,P2,Q1,Q2,data){
  
  #K=(p2+1)*(q2+1)*(P2+1)*(Q2+1)
  
  temp=c()
  d=1
  D=1
  s=20
  
  i=1
  temp= data.frame()
  ls=matrix(rep(NA,9*35),nrow=35)
  
  
  for (p in p1:p2)
  {
    for(q in q1:q2)
    {
      for(P in P1:P2)
      {
        for(Q in Q1:Q2)
        {
          if(p+d+q+P+D+Q<=9)
          {
            
            model<- Arima(data,order=c(p-1,d,q-1),seasonal=c(P-1,D,Q-1))
            ls[i,]= c(p-1,d,q-1,P-1,D,Q-1,model$aic,model$bic,model$aicc)
            i=i+1
            #print(i)
            
          }
          
        }
      }
    }
    
  }
  
  
  temp= as.data.frame(ls)
  names(temp)= c("p","d","q","P","D","Q","AIC","BIC","AICc")
  
  temp
  
}

```



```{r}
# q=0,1,2,3; Q=1,2 and PACF plot: p=0,1,2; P=1,2, D=1 and d=0,1
output=SARIMA.c(p1=1,p2=4,q1=1,q2=2,P1=1,P2=3,Q1=1,Q2=2,data=temp_month)
#output

knitr::kable(output)
```
```{r}
output[which.min(output$AIC),] 
output[which.min(output$BIC),]
output[which.min(output$AICc),]

set.seed(123)
model_output <- capture.output(sarima(temp_month, 1,1,1,1,1,0,12))
```

```{r}
fit <- Arima(temp_month, order=c(1,1,1), seasonal=c(1,1,0))
summary(fit)
```

```{r}
fit <- Arima(temp_month, order=c(1,1,1), seasonal=c(1,1,0))
summary(fit)

# Forecasting
fit %>% forecast(h=36) %>% autoplot(main="Forecast for Monthly Temperature in DC",ylab="Tempreature") #next 3 years
sarima.for(temp_month, 36, 1,1,1,1,1,0,12)
```

```{r, message=FALSE, warning=FALSE}

# Compare with benchmark methods
f1 <- meanf(d1, h=20) 

checkresiduals(f1)

f2 <- naive(d1, h=20) 

checkresiduals(f2)

f3 <- rwf(d1,drift=TRUE, h=20) 

checkresiduals(f3)

```


```{r}
# Do a seasonal cross validation using 1 step ahead Forecasting
k <- length(fit)
n <- length(temp_month)
n-k
```

```{r}
i=1
err = c()

for(i in 1:(n-k))
{
  xtrain <- temp_month[1:(k-1)+i] #observations from 1 to 18
  xtest <- temp_month[k+i] #18th observation as the test set
  
  fit <- Arima(xtrain, order=c(1,1,1),include.drift=FALSE, method="ML")
  fcast <- forecast(fit, h=1)
  
  #capture error for each iteration
  # This is mean absolute error
  err = c(err, abs(fcast$mean-xtest)) 

  # This is mean squared error
  err2 = c(err, (fcast$mean-xtest)^2)

}

(MAE=mean(err)) # This is mean absolute error
(RMAE=mean(err2)) # This is mean squared error

```

```{r}
# Do a seasonal cross validation using 12 steps ahead forecasts
k <- length(fit)
n <- length(log(temp_month))
n-k

mae <- matrix(NA,n-k,12)
st <- tsp(log(temp_month))[1]+(k-2)/12

head(log(temp_month))

```

```{r,warning=FALSE}
for(i in 1:(n-k))
{
  #xtrain <- window(a10, start=st+(i-k+1)/12, end=st+i/12)
  xtrain <- window(log(temp_month), end=st + i/12)
  xtest <- window(log(temp_month), start=st + (i+1)/12, end=st + (i+12)/12) #forecasting for 12 steps; so it goes from (i+1)/12 to (i+12)/12
  
  fit <- Arima(xtrain, order=c(1,1,1), seasonal=list(order=c(0,1,1), period=12),
                include.drift=FALSE, lambda=0, method="ML")
  fcast <- forecast(fit, h=12)
  
  mae[i,1:length(xtest)] <- abs(fcast$mean-xtest)

}

plot(1:12, colMeans(mae,na.rm=TRUE), type="l", col=2, xlab="horizon", ylab="MAE",main = "MAE For 12 Steps Ahead Forecast")
```




